Big Ideas in Sports Analytics and Statistical Tools for their Investigation

### 2022-06-01 sports_analytics.Rmd
added Handbook references
added BTM computations

Numerous papers (see @lopez2020bigger for details) have used the analysis of the expected number of points to improve fourth down strategy.

@@

@@

@@

```{r btm, echo=TRUE}
library(Lahman)
library(BradleyTerry2)
# show BTM for Pirates vs. Yankees in 1960
# or Red Sox vs. Yankees in 2003/2004
```

@@

@Glickman1998 state-space model

@@

>>>>

Numerous papers (see @lopez2020bigger for details) have used the analysis of the expected number of points to improve fourth down strategy. 
@goldner2017situational uses a similar framework.

@@

See also @bornn2017studying.

@@

@goldner2017situational

@@

```{r btm, echo=TRUE}
library(tidyverse)
library(BradleyTerry2)
# show BTM for Pirates vs. Yankees in 1960
# or Red Sox vs. Yankees in 2003/2004
library(retrosheet)
games <- getRetrosheet("game", 2004)

bt_games <- games |>
  group_by(HmTm, VisTm) |>
  summarize(
    home_w = sum(HmRuns > VisRuns),
    away_w = sum(VisRuns > HmRuns),
  ) |>
  mutate(
    HmTm = factor(HmTm),
    VisTm = factor(VisTm, levels = levels(HmTm))
  )

btm <- BTm(
  outcome = cbind(home_w, away_w), 
  player1 = HmTm, 
  player2 = VisTm,
  data = bt_games, 
  id = "team"
)
btm

bt_games2 <- bt_games |>
  mutate(
    HmTm = data.frame(team = HmTm, at_home = 1),
    VisTm = data.frame(team = VisTm, at_home = 0)
  )

btm2 <- btm |>
  update(formula = ~ team + at_home, data = bt_games2)
btm2
```

```{r}
btm |>
  broom::tidy() |>
  arrange(desc(estimate))
btm2 |>
  broom::tidy() |>
  arrange(desc(estimate))
```

@@

@Glickman1998 state-space model. 
See also @glickman2017estimating.

@@

In soccer, @koning2017rating
